Notes for Week 2/28/20 - 3/5/20

### 2020-03-08 notes/weekly_notes/3.5.20.NotesForWeek.tex
updated weekly notes with tighter grid approximation error bound

\section{Sobolev testing with fixed design.}
For now, we confine our attention to the periodic Sobolev space $H_{\textrm{per}}^{s}([0,1]^d)$ which consists of all those functions $f \in \Leb^2([0,1]^d)$ whose tensor product Fourier coefficients decay at a sufficiently fast rate. When $d = 1$ and $x \in [0,1]$, let
\begin{align*}
\phi_1(x) & = 1 \\
\phi_{2k}(x) & = \cos(2k\pi x),~~ k \in \mathbb{N} \\
\phi_{2k + 1}(x) & = \sin(2k\pi x),~~ k \in \mathbb{N}
\end{align*}
be the trigonometric basis over $[0,1]$. The tensor product Fourier basis is defined on $[0,1]^d$ as
\begin{equation*}
\phi_k(x) = \prod_{i = 1}^{d} \phi_{k_i}(x_i),
\end{equation*}
the Sobolev space $H_{\textrm{per}}^{s}([0,1]^d)$ as 
\begin{equation*}
H^{\textrm{per}}([0,1]^d) = \set{\sum_{k \in \mathbb{N}^d} \theta_k \phi_k^{(d)}: \sum_{k \in \mathbb{N}^d} \theta_k^2 \abs{k}^{2s} < \infty},
\end{equation*}
and the unit ball $H_{\textrm{per}}^{s}([0,1]^d, L)$ as 
\begin{equation*}
H^{\textrm{per}}([0,1]^d;L) = \set{\sum_{k \in \mathbb{N}^d} \theta_k \phi_k^{(d)}: \sum_{k \in \mathbb{N}^d} \theta_k^2 \abs{k}^{2s} \leq L},
\end{equation*}

@@

It is possible to improve on Lemma~\ref{lem:grid_sobolev_approximation_error} under \textcolor{red}{(relatively mild)} additional assumptions. Specifically, we assume the existence of $\varepsilon_j$ satisfying $j \varepsilon_j$ is monotone non-increasing in $j$, such that for every $f \in H_{\textrm{per}}^{s}([0,1]^d)$, $f = \sum_{k \in \mathbb{N}^d} \theta_k \phi_k$,
\begin{equation}
\label{asmp:fourier_coefficient_decay}
\abs{\theta_k} \leq \prod_{i = 1}^{d} \varepsilon_{k_i}~~\textrm{for all $k \in \mathbb{N}^d$},~~ \sum_{j = 1}^{\infty} \varepsilon_j = C,~~ \sum_{j = 1}^{\infty} \varepsilon_j^2 \leq C \biggl(\sum_{k \in \mathbb{N}^d} \theta_k^2\biggr)^{1/d}
\end{equation}
for some constant $C < \infty$.

\textcolor{red}{(TODO:)} {Unfortunately these additional assumptions are not mild at all, but instead relatively close to assuming the function is $s$-Holder.}

@@

Now we will attempt to find a sharper bound on discretization error than Lemma~\ref{lem:grid_sobolev_approximation_error}, but that requires less severe assumptions than Lemma~\ref{lem:grid_sobolev_approximation_error_2}.
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_3}
	For any $2s > d$ and $f \in H_{\textrm{per}}^s([0,1]^d;L)$, 
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_{n}^2 \leq c n^{1/2 - s/d} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation*}
\end{lemma}

@@

\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_4}
	Suppose $2s < d$ and $f = \sum_{k \in \mathbb{N}^k} \theta_k \varphi_k$ satisfies $f \in \wt{H}_{\textrm{per}}^{s}([0,1]^d;L)$. Then
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c n^{1/2 - s/d} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation*}
\end{lemma}

@@

@@

	The same statement holds true when replacing $\phi_k$ by $\varphi_k$ and $\phi_l$ by $\varphi_l$, and using symmetric grid points $\wb{X}_i$.
\end{lemma}
\begin{proof}
	We prove only the statement involving $\varphi_k$ and $\varphi_l$. Note that for $k = 1,\ldots,n$, and $i = 1,\ldots,n$,
	\begin{equation*}
	v_{k,i} = \frac{1}{\sqrt{n}}\varphi_k(\wb{x}_i)
	\end{equation*}
	and $\Delta_{kl}^{(n)} = 0$ for $1 \leq k,l \leq n$ immediately follows.

	By restricting the basis $\varphi_k$ to the nodes $\wb{x}_i$ we induce periodicity. In particular, for $k = mn + (j + 1)$ where $m$ and $j$ are non-negative integers and $0 \leq j < n$, we have
	\begin{equation*}
	\varphi_k(\wb{x}_i) \propto
	\begin{cases*}
	\varphi_{j + 1}(\wb{x}_i),& ~~\textrm{if $m \mod 2 = 0$} \\
	\varphi_{n - j}(\wb{x}_i),& ~~\textrm{if $m \mod 2 = 1$}
	\end{cases*}
	\end{equation*}
	Writing $l = m'n + j' + 1$, we therefore see that $\dotp{\varphi_k}{\varphi_l}_n \in \{-1,+1\}$ if and only if (a) $m \mod 2 = m' \mod 2$ and $j = j'$, or (b) $m \mod 2 \neq m' \mod 2$ and $j + 1 = n - j'$; otherwise, $\dotp{\varphi_k}{\varphi_l}_n = 0$. Condition (a) is equivalent to $|[k/2] - [l/2]| = mn$ for some $m \in \mathbb{N}$, and Condition (b) is equivalent to $[k/2] + [m/2] = mn$ for some $m \in \Nbb$, thus proving the claim.

>>>>

\section{Sobolev testing with fixed design.}
For now, we confine our attention to the periodic Sobolev space $H_{\textrm{per}}^{s}([0,1]^d)$ which consists of all those functions $f \in \Leb^2([0,1]^d)$ whose tensor product Fourier coefficients decay at a sufficiently fast rate. When $d = 1$ and $x \in [0,1]$, let
\begin{align*}
\phi_1(x) & = 1 \\
\phi_{2k}(x) & = \sqrt{2}\cos(2k\pi x),~~ k \in \mathbb{N} \\
\phi_{2k + 1}(x) & = \sqrt{2}\sin(2k\pi x),~~ k \in \mathbb{N}
\end{align*}
be the trigonometric basis over $[0,1]$. The tensor product Fourier basis is defined on $[0,1]^d$ as
\begin{equation*}
\phi_k(x) = \prod_{i = 1}^{d} \phi_{k_i}(x_i),
\end{equation*}
the Sobolev space $H_{\textrm{per}}^{s}([0,1]^d)$ as 
\begin{equation*}
H^{\textrm{per}}([0,1]^d) = \set{\sum_{k \in \mathbb{N}^d} \theta_k \phi_k^{(d)}: \sum_{k \in \mathbb{N}^d} \theta_k^2 \abs{k}^{2s} < \infty},
\end{equation*}
and the unit ball $H_{\textrm{per}}^{s}([0,1]^d, L)$ as 
\begin{equation*}
H^{\textrm{per}}([0,1]^d;L) = \set{\sum_{k \in \mathbb{N}^d} \theta_k \phi_k^{(d)}: \sum_{k \in \mathbb{N}^d} \theta_k^2 \abs{k}^{2s} \leq L},
\end{equation*}

@@

It is possible to improve on Lemma~\ref{lem:grid_sobolev_approximation_error} under additional assumptions. Specifically, we assume the existence of $\varepsilon_j$ satisfying $j \varepsilon_j$ is monotone non-increasing in $j$, such that for every $f \in H_{\textrm{per}}^{s}([0,1]^d)$, $f = \sum_{k \in \mathbb{N}^d} \theta_k \phi_k$,
\begin{equation}
\label{asmp:fourier_coefficient_decay}
\abs{\theta_k} \leq \prod_{i = 1}^{d} \varepsilon_{k_i}~~\textrm{for all $k \in \mathbb{N}^d$},~~ \sum_{j = 1}^{\infty} \varepsilon_j = C,~~ \sum_{j = 1}^{\infty} \varepsilon_j^2 \leq C \biggl(\sum_{k \in \mathbb{N}^d} \theta_k^2\biggr)^{1/d}
\end{equation}
for some constant $C < \infty$.

@@

Unfortunately the additional assumption~\eqref{asmp:fourier_coefficient_decay} is not mild at all, but instead relatively close to assuming the function is $s$-Holder.
Finally, we state a sharper bound on discretization error than Lemma~\ref{lem:grid_sobolev_approximation_error}, but that requires less severe assumptions than Lemma~\ref{lem:grid_sobolev_approximation_error_2}.
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_3}
	For any $2s > d$ and $f \in H_{\textrm{per}}^s([0,1]^d;L)$, 
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_{n}^2 \leq c n^{1/2 - s/d} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation*}
\end{lemma}

@@

\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_4}
	Suppose $2s < d$ and $f = \sum_{k \in \mathbb{N}^k} \theta_k \varphi_k$ satisfies $f \in \wt{H}^{s}([0,1]^d;L)$. Then
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c n^{1/2 - s/d} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation*}
\end{lemma}

So we've done a lot of interesting work, but we only obtained a sufficient bound on the approximation error when $n^{1/2 - s/d} = o(\norm{f}_2)$, which is still not good enough for our purposes. In the following Lemma, we state a tighter bound on approximation error when $d = 1$.
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_5}
	Suppose $2s > 1$. Then there exists a constant $c$ which depends only on $s$ such that for any $f = \sum_{k = 1}^{\infty} \theta_k \varphi_k \in \wt{H}_{\textrm{per}}^{s}([0,1]^d;L)$, the following bound holds:
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_1}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s} + \sum_{k = \kappa + 1}^{\infty} \theta_k^2
	\end{equation*} 
\end{lemma}

We extend the results of Lemma~\ref{lem:grid_sobolev_approximation_error_5} to hold for all $d \geq 1$ in Lemma~\ref{lem:grid_sobolev_approximation_error_6}. Although Lemma~\ref{lem:grid_sobolev_approximation_error_6} strictly dominates Lemma~\ref{lem:grid_sobolev_approximation_error_5}, we keep the latter around since its proof is somewhat easier to follow. 
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_6}
	Suppose $f = \sum_{k \in \Nbb^d} \theta_k \varphi_k$ satisfies $f \in \wt{H}^s([0,1]^d;L)$ for some $2s > d$. Then, there exists a constant $c$ which depends only on $s$ and $d$ such that
	\begin{equation}
	\label{eqn:grid_sobolev_approximation_error_6}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s/d} + \sum_{k \in \Nbb \setminus [K]^d} \theta_k^2
	\end{equation}
\end{lemma}

@@

\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_5}}
Recall that the eigenvectors of the grid graph $v_k := v_k(\wb{G})$ satisfy
\begin{equation*}
v_{k}(\wb{x}_i) = \frac{1}{\sqrt{n}} \varphi_{k}(\wb{x}_i)
\end{equation*}
for each $i \in [n]$ and $k \in \mathbb{N}$. Therefore we may rewrite $\Pi_{\kappa,\wb{G}}$ as a function of the cosine Fourier basis,
\begin{align*}
\Pi_{\kappa,\wb{G}}(f) & = \frac{1}{n} \sum_{k = 1}^{\kappa} \Bigl\{\sum_{i \in [n]} f(\wb{x}_i) \varphi_k(\wb{x}_i)\Bigr\} \varphi_k \\
& :=  \sum_{k = 1}^{\kappa} \wt{\theta}_k \varphi_k
\end{align*}
where we interpret the right hand side as a function over $\wb{X}$. Since $(\varphi_k)$ is an $L_2(\wb{X})$ orthonormal sequence, we obtain
\begin{align*}
\norm{\Pi_{\kappa,\wb{G}}(f)}_n^2 & = \sum_{k = 1}^{\kappa} \wt{\theta}_k^2 \\ 
& = \sum_{k = 1}^{\kappa} {\theta}_k^2 + \sum_{k = 1}^{\kappa} \wt{\theta}_k^2 - \theta_k^2 \\
& = \norm{f}_{\Leb^2([0,1]^d)}^2 - \sum_{k = \kappa + 1}^{\infty} \theta_k^2 + \sum_{k = 1}^{\kappa} \wt{\theta}_k^2 - \theta_k^2
\end{align*}
or equivalently
\begin{equation*}
\norm{f}_{\Leb^2([0,1]^d)}^2 - \norm{\Pi_{\kappa,\wb{G}}(f)}_n^2 = \sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 + \sum_{k = \kappa + 1}^{\infty} \theta_k^2
\end{equation*}
The second term in the above equation is exactly the second term in~\eqref{eqn:grid_sobolev_approximation_error}, and so we focus our attention on the first term, which represents discretization error. Applying the parallelogram law and the Cauchy-Schwarz inequality gives
\begin{equation}
\label{eqn:grid_sobolev_approximation_error5_pf2}
\begin{aligned}
\sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 & \leq \sum_{k = 1}^{\kappa} \max\Bigl\{(\theta_k + \wt{\theta}_k)(\theta_k - \wt{\theta}_k),0\Bigr\} \\
& \leq \left(\sum_{k = 1}^{\kappa} \max\Bigl\{(\theta_k + \wt{\theta}_k)^2,0\Bigr\}\right)^{1/2} \left(\sum_{k = 1}^{\kappa} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} \\
& \leq \sqrt{2} \norm{f}_{\Leb^2([0,1])} \left(\sum_{k = 1}^{\kappa} (\theta_k - \wt{\theta}_k)^2\right)^{1/2}
\end{aligned}
\end{equation}

It is known that the error $\theta_k - \wt{\theta}_k$ is due to the aliasing phenomenon, where high frequency sinusoidal waves agree with low frequency waves at evenly spaced points. This phenomenon has a very precise periodic nature -- expressed in Lemma~\ref{lem:alias}--and we use this Lemma to get the upper bound
\begin{align*}
\abs{{\theta}_k - \wt{\theta}_k} & = \abs{\sum_{\ell = 1}^{\infty} \theta_{\ell} \frac{1}{n} \sum_{i = 1}^{n} \varphi_{\ell}(\wb{x}_i) \varphi_{k}(\wb{x}_i)} \\
& = \sum_{m = 1}^{\infty} \abs{\theta_{2mn + k}} + \sum_{m = 0}^{\infty}\abs{\theta_{2(m + 1)n - (k - 1)}}.
\end{align*}
Since $\theta$ belongs to the Sobolev ellipsoid $\Theta(s,L)$, by applying Cauchy-Schwarz we can convert the above two tail sums to bounds which explicitly depend on $n$. Applying this logic to the first tail sum gives us
\begin{align}
\sum_{m = 1}^{\infty} \abs{\theta_{2mn + k}} & \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k})^{-2}\biggr)^{1/2}\nonumber \\
& \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl( \sum_{m = 1}^{\infty} (2mn)^{-2s}\biggr)^{1/2} \nonumber \\
& \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl( \Bigl(1 + \int_{1}^{\infty} x^{-2s} \,dx \Bigr) \biggr)^{1/2} (2n)^{-s}   \nonumber \\
& = \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl( \Bigl(1 + \frac{1}{2s - 1} \Bigr) \biggr)^{1/2} (2n)^{-s}\label{eqn:grid_sobolev_approximation_error5_pf3}
\end{align}
where we have used the fact that the Riemann sum of a monotone non-increasing function evaluated at right end points is always no greater than the corresponding integral, and the last equality holds since $2s > 1$. By using essentially equivalent reasoning, we obtain a comparable bound on the second tail sum
\begin{equation*}
\sum_{m = 0}^{\infty}\abs{\theta_{2(m + 1)n - (k - 1)}} \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2(m + 1)n - (k - 1)}\theta_{2(m + 1)n - (k - 1)}\bigr)^2\biggr)^{1/2} \biggl( \Bigl(2 + \frac{1}{2s - 1} \Bigr) \biggr)^{1/2} (2n)^{-s}
\end{equation*}
Thus,
\begin{align}
\sum_{k = 1}^{\kappa} \bigl(\theta_k - \wt{\theta}_k\bigr)^2 & \leq c(s) n^{-2s}\sum_{k = 1}^{\kappa} \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2 + \sum_{m = 0}^{\infty}\bigl(a_{2(m + 1)n - (k - 1)}\theta_{2(m + 1)n - k + 1} \bigr)^2\biggr) \nonumber \\
& \leq c(s) n^{-2s} \sum_{j = n + 1}^{\infty} \bigl(\theta_j^2 a_j^2)
\nonumber \\
& \leq L c(s) n^{-2s}. \label{eqn:grid_sobolev_approximation_error5_pf4}
\end{align}
where the second inequality in the preceding display follows since for any $(m,k) \neq (m',k') \in \mathbb{N} \times [n]$,
\begin{equation*}
2mn + k \neq 2m'n + k',~~ 2(m + 1)n - k + 1 \neq 2(m' + 1)n - k' + 1, ~~\textrm{and}~~  2mn + k \neq 2(m' + 1)n - k' + 1
\end{equation*}
We plug~\eqref{eqn:grid_sobolev_approximation_error5_pf4} back in to~\eqref{eqn:grid_sobolev_approximation_error5_pf2} to obtain the claim of the Lemma.

\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_6}}
Recall that the eigenvectors of the grid graph $v_k := v_k(\wb{G}_d)$ satisfy
\begin{equation*}
v_{k}(\wb{x}_i) = \frac{1}{\sqrt{n}} \varphi_{k}(\wb{x}_i)
\end{equation*}
for each $i \in [N]^d$ and $k \in \mathbb{N}^d$. Therefore we may rewrite $\Pi_{\kappa,\wb{G}_d}$ as a function of the tensor product cosine Fourier basis,
\begin{align*}
\Pi_{\kappa,\wb{G}_d}(f) & = \frac{1}{n} \sum_{k \in [K]^d} \Bigl\{\sum_{i \in [N]^d} f(\wb{x}_i) \varphi_k(\wb{x}_i)\Bigr\} \varphi_k \\
& :=  \sum_{k \in [K]^d} \wt{\theta}_k \varphi_k
\end{align*}
where we interpret the right hand side as a function over $\wb{X}$. Since $(\varphi_k)$ is an $L_2(\wb{X})$ orthonormal sequence, we obtain
\begin{align*}
\norm{\Pi_{\kappa,\wb{G}_d}(f)}_n^2 & = \sum_{k \in [K]^d} \wt{\theta}_k^2 \\ 
& = \sum_{k \in [K]^d} {\theta}_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2 \\
& = \norm{f}_{\Leb^2([0,1]^d)}^2 - \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2
\end{align*}
or equivalently
\begin{equation*}
\norm{f}_{\Leb^2([0,1]^d)}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_n^2 = \sum_{k \in [K]^d} \theta_k^2 - \wt{\theta}_k^2 + \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2
\end{equation*}
The second term in the above equation is exactly the second term in~\eqref{eqn:grid_sobolev_approximation_error_6}, and so we focus our attention on the first term, which represents discretization error. Applying the parallelogram law and the Cauchy-Schwarz inequality gives
\begin{equation}
\label{eqn:grid_sobolev_approximation_error6_pf1}
\begin{aligned}
\sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 & \leq \sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)(\theta_k - \wt{\theta}_k),0\Bigr\} \\
& \leq \left(\sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)^2,0\Bigr\}\right)^{1/2} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} \\
& \leq \sqrt{2} \norm{f}_{\Leb^2([0,1])} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2}
\end{aligned}
\end{equation}

It is known that the error $\theta_k - \wt{\theta}_k$ is due to the aliasing phenomenon, where high frequency sinusoidal waves agree with low frequency waves at evenly spaced points. This phenomenon has a very precise periodic nature, as expressed in Lemma~\ref{lem:alias}, which allows us to rewrite $\wt{\theta}_k$. Note that for every $\ell \in \mathbb{N}^d$, there exists exactly one $m \in \mathbb{N}_0^d$, $o \in [N]^d$ and $b \in \{0,1\}^d$ such that $\ell = 2mN + (1 - b)o + b(-(o - 1))$ (where for $a,b \in \Reals^d$ we denote $ab := (a_1b_1,\ldots,a_db_d)$). Adopting the convention $\theta_{\ell} := 0$ if $\ell$ contains a negative index, we have
\wt{\theta}_k & = \frac{1}{n} \sum_{i \in [N]^d} f(\wb{x}_i) \varphi_k(\wb{x}_i) \\
& = \frac{1}{n} \sum_{i \in [N]^d} \Bigl(\sum_{\ell \in \Nbb^d} \theta_{\ell} \varphi_{\ell}(\wb{x}_i) \Bigr) \varphi_k(\wb{x}_i) \\
& = \sum_{\ell \in \mathbb{N}^d} \theta_{\ell} \Bigl(\frac{1}{n} \sum_{i \in [N]^d} \varphi_{\ell}(\wb{x}_i) \varphi_{k}(\wb{x}_i)\Bigr) \\
& = \sum_{m \in \mathbb{N}_0^d} \sum_{o \in [N]^d}\sum_{b \in \{0,1\}^d} \theta_{I_N(m,b,o)}  \Bigl(\frac{1}{n} \sum_{i \in [N]^d} \varphi_{I_N(m,b,o)}(\wb{x}_i) \varphi_{k}(\wb{x}_i)\Bigr) \\
& = \sum_{m \in \mathbb{N}_0^d} \sum_{b \in \{0,1\}^d} \theta_{I_N(m,b,k)} 
where the last equality follows from Lemma~\ref{lem:alias}. By applying the Cauchy-Schwarz inequality, and noting that $I_N(m,b,k) \geq mN$ (where inequality between two vectors is interpreted entrywise) we get
\abs{\wt{\theta}_k - \theta_k} & \leq \sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \abs{\theta_{I_N(m,b,k)}} \\
& \leq \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \biggr)^{1/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(a_{I_N(m,b,k)}\Bigr)^{-2} \biggr)^{1/2} \\
& \leq 2^{d/2}N^{-s } \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \biggr)^{1/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}}\sum_{j = 1}^{d} m_j^{-2s} \biggr)^{1/2} 
Letting $c(d,s)$ be a finite constant which may change from line to line but can depend only on $d$ and $s$, and letting $\Nbb_2 = \Nbb \setminus \{1\}$, we can upper bound the second sum in the previous display,
\sum_{m \in \Nbb_0^d \setminus \{0\}}\sum_{j = 1}^{d} m_j^{-2s} & \overset{(i)}{\leq} d^{2(s + 1)} \sum_{m \in \Nbb_0^d \setminus \{0\}} \Bigl(\sum_{j = 1}^{d} m_j^{-2}\Bigr)^s \\
& \overset{(ii)}{\leq} d^{2(s + 1)} \sum_{\wt{d} = 1}^{d} 2^{\wt{d}} \frac{d!}{\wt{d}!} \sum_{m \in \Nbb_2^{\wt{d}}} \Bigl(\sum_{j = 1}^{d} m_j^{-2}\Bigr)^s \\
& \overset{(iii)}{\leq} d^{2(s + 1)} \sum_{\wt{d} = 1}^{d} 2^{\wt{d}} \frac{d!}{\wt{d}!} \int_{[1,\infty)^{\wt{d}}} \Bigl(\sum_{j = 1}^{\wt{d}} x_j^{-2}\Bigr)^{s} \,dx_1 \ldots \,dx_{\wt{d}} \\
& \leq \sum_{\wt{d} = 1}^{d} c(d,s) \int_{\sqrt{\wt{d}}}^{\infty} r^{-2s + d - 1} \,dr \\
& \overset{(iv)}{\leq} c(d,s) < \infty.
where $(i)$ is an equality when $d = 1$, and otherwise follows from Jensen's inequality when $d \geq 2$ and $s \geq 1$; $(ii)$ follows from specially considering the edge cases where $m_j = 0$ or $m_j = 1$ for $j = \wt{d} + 1,\ldots, d$; $(iii)$ uses the fact that the Riemann sum of a monotone non-increasing function evaluated at right end points is always no greater than the corresponding integral; and $(iv)$ follows since $2s > d$. We are now in a position to conclude that
\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2 & \leq c(d,s) N^{-2s} \sum_{k \in [K]^d} \sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \nonumber \\
& \leq c(d,s) N^{-2s} \sum_{k \in \Nbb^d \setminus [N]^d} \theta_k^2 a_k^2 \nonumber \\
& \leq L c(d,s) N^{-2s} \label{eqn:grid_sobolev_approximation_error6_pf2}
where the second inequality in the preceding display follows since for any $(m,k,b) \neq (m',k',b')$, $I_N(m,b,k) \neq I_N(m',b',k')$. We plug~\eqref{eqn:grid_sobolev_approximation_error6_pf2} back in to~\eqref{eqn:grid_sobolev_approximation_error6_pf1} to obtain the claim of the Lemma.

@@

	The same statement holds true when replacing $\phi_k$ by $\varphi_k$ and $\phi_l$ by $\varphi_l$, and using symmetric grid points $\wb{X}_i$.
\end{lemma}

The following Lemma is similar to~\eqref{lem:polyak90}, but tweaked to be more useful for our setting. It applies to the elements in the cosine basis of period $2$, rather than elements in the Fourier basis of period $1$, and is stated slightly differently.
\begin{lemma}
	\label{lem:alias}
	Let $k \in [n]$ and let $\ell \in \Nbb$. There exists a unique $m \in \Nbb_{0}$, $j \in [n]$ and $b \in \{0,1\}$ such that $\ell = 2mn + (1 - b)j - b(j - 1)$. Then
	\begin{equation*}
	\frac{1}{n}\sum_{i = 1}^{n} \varphi_{k}(\wb{x}_i)\varphi_{\ell}(\wb{x}_i) =  
	\begin{cases*}
	1,& ~~ \textrm{if $j = k$} \\
	0,& ~~ \textrm{otherwise.}
	\end{cases*}
	\end{equation*}
\end{lemma}

\begin{proof}
	Note that for $k = 1,\ldots,n$, and $i = 1,\ldots,n$,
	\begin{equation*}
	v_{k,i} = \frac{1}{\sqrt{n}}\varphi_k(\wb{x}_i).
	\end{equation*}
	Additionally, it is easy to check that
	\begin{equation*}
	\varphi_{\ell}(\wb{x}_i) \propto
	\begin{cases*}
	\varphi_{j}(\wb{x}_i),& ~~\textrm{if $j \in [n]$} \\
	\varphi_{-(j - 1)}(\wb{x}_i),& ~~\textrm{if $j \in [-(n - 1)]$} \\
	\varphi_{1}(\wb{x}_i),& ~~\textrm{if $j = 0$.}
	\end{cases*}
	\end{equation*}
	The claim follows since $(v_k)$ is orthonormal in empirical norm.	
\end{proof}
